---
title: "Bayes_Rule"
author: "Peiyuan Zhu"
date: "2023-11-02"
output: rmarkdown::html_vignette
# output: word_document
vignette: >
  %\VignetteIndexEntry{Bayes_Rule}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# devtools::load_all(".") # only used in place of dst when testing with R-devel
library(dst) 
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In Mathematical Theory of Evidence Glenn Shafer talked about how Dempster's rule of combination generalizes Bayesian conditioning. In this document we investigate numerically how a simple Bayesian model can be encoded into the language of belief function. 

Recall the Bayes Rule of conditioning in simple terms:

$$P(H|E) = \dfrac{P(H) \cdot P(E|H)} {P(E)}$$
Let's see how this is translated in the belief functions setup.

#  1. Simple Bayes Example

In particular, the Bayesian belief functions concentrates their masses on the singletons only, unlike more general basic mass assignment functions. For instance, in a frame $\Theta=\{a,b,c\}$, basic mass assignment $m(\{a\})=0.2$, $m(\{b\})=0.3$ and $m(\{c\})=0.5$ defines a Bayesian belief function. 

In the Bayesian language, this is the prior distribution $P(H)$. Function *bca* is used to set the distribution of *H*.

```{r bpa1 definition,  echo = FALSE, warning=FALSE}
Theta<-matrix(c(1,0,0,0,1,0,0,0,1,1,1,1), nrow = 4, byrow = TRUE)
H <- bca(tt=matrix(c(1,0,0,0,1,0,0,0,1), nrow = 3, byrow = TRUE), m = c(0.2, 0.3, 0.5), cnames = c("a", "b", "c"), idvar = 1)
cat("The prior distribution H","\n")
bcaPrint(H)
# round(belplau(H, h=Theta), digits = 3)
```
The law of conditional probability is a special case of Dempster's rule of combination that all the masses focus on the event is conditioned. For instance, basic mass assignment focuses all the masses on subset $E =\{b,c\}$. Hence, using function *bca*, we set $m(\{b,c\})=1$. 

```{r bpa2 definition,  echo = FALSE, warning=FALSE}
bpa2 <- bca(tt=rbind(diag(x=1, nrow=3), matrix(c(0,1,1,1,1,1), nrow=2, byrow = TRUE)), m = c(0,0,0,1,0), cnames = c("a", "b", "c"), idvar = 1)
Event <-  addTobca(bpa2, tt = diag(x=1, nrow = 3))
cat("Setting an Event E = {b,c} with mass = 1","\n")
bcaPrint(Event)
```

Now we set the computation of Bayes's Theorem in motion. 

In a first step, we use function *dsrwon* to combine our two basic mass assignments H and Event. The non-normalized Dempster Rule of combination gives a mass distribution *H_Event* composed of two parts:

1. the distribution of the product $P(H) \cdot P(E|H)$ on $\Theta$;
2. a mass allotted to the empty set $m(\varnothing)$.

```{r H_Event Dempster_rule1,  echo = FALSE, warning=FALSE}
H_Event <- dsrwon(H, bpa2)
cat("The combination of H and Event E","\n")
bcaPrint(H_Event)
```
It turns out that we can obtain the marginal $P(E)$ from $m(\varnothing)$:
$$P(E) = 1 - m(\varnothing)$$.

Hence, $P(E)$ is nothing else than the normalization constant of Dempster's rule of combination. 

In our second step of computation we us function *nzdsr*, to apply the normalization constant to distribution *H_Event*, which gives the posterior distribution $P(H|E)$
 
```{r H_Event Dempster_rule2,  echo = FALSE, warning=FALSE}
H_given_E <- nzdsr(H_Event)
cat("The posterior distribution P(H|E)","\n")
bcaPrint(H_given_E)
```

Note that *H_given_E* is defined only on singletons and the mass allocated to $\Theta$ is zero. Hence $bel(\cdot) = P(\cdot) = Pl(\cdot)$, as shown by the following table.

 
```{r H_Event Dempster_rule3,  echo = FALSE, warning=FALSE}
round(belplau(H_given_E, h=Theta), digits = 3)
```

# 2. Example with  two variables

In the first example, the conditioning event was a subset of the frame $\Theta$ of variable *H*. We now show the computation of Bayes's rule of conditioning by Dempster's Rule in the case of two variables. 

Let's say we have the variable H defined on $\Theta = \{a, b, c\}$ as before.
```{r bpa1_copy,  echo = FALSE, warning=FALSE}
Theta<-matrix(c(1,0,0,0,1,0,0,0,1,1,1,1), nrow = 4, byrow = TRUE)
X <- bca(tt=matrix(c(1,0,0,0,1,0,0,0,1), nrow = 3, byrow = TRUE), m = c(0.2, 0.3, 0.5), cnames = c("a", "b", "c"), idvar = 1, varnames = "x")
cat("The prior distribution","\n")
bcaPrint(X)
```
let's add a second variable E with three outcomes $\Lambda =\{d, e, f\}$ .

$P(\{d|a\})=0.1$, $P(\{d|b\})=0.2$ and $P(\{d|c\})=0.7$. 

This distribution will be encoded in the product space $\Theta \times \Lambda$ by setting 

$m(\{a,d\}) = 0.1$; $m(\{b,d\}) = 0.2$; $m(\{c,d\}) = 0.7$

We now do this using function *bcaRel*.

```{r relation,  echo = FALSE, warning=FALSE}
# bpa4 <- bca(tt=matrix(c(1,0,0,0,1,0,0,0,1), nrow = 3, byrow = TRUE), m = c(1, 0, 0), cnames = c("d", "e", "f"), idvar = 4, varnames = "y")
# bcaPrint(bpa4)
#
cat("Specify information on variables, description matrix and mass vector","\n")
inforvar_EX <- matrix(c(1,4,3,3), ncol = 2,  dimnames = list(NULL, c("varnb", "size")) )
cat("Identifying variables and frames","\n")
inforvar_EX
cat("Note that variables numbers must be in increasing order","\n")
# 
tt_EX <- matrix(c(1,0,0,1,0,0,
                   0,1,0,1,0,0,
                   0,0,1,1,0,0,
                   1,1,1,1,1,1), ncol = 6, byrow = TRUE, dimnames = list(NULL, c("a", "b", "c", "d", "e", "f")))
cat("The description matrix of the relation between X and E","\n")
tt_EX
cat("Note Columns of matrix must follow variables ordering. ","\n")
#
spec_EX <-  matrix(c(1:4, 0.1, 0.2, 0.7, 0 ), ncol = 2, dimnames = list(NULL, c("specnb", "mass")))
cat("Mass specifications","\n")
spec_EX
# 
rel_EX <- bcaRel(tt = tt_EX, spec = spec_EX, infovar = inforvar_EX, varnames = c("x", "y"), relnb = 1)
cat("The relation between Evidence E and X","\n")
bcaPrint(rel_EX)
```

Now we combine Prior $P(X)$ with rel_EX. But first, we need to extent *X* to the space $\Theta \times \Lambda$.

```{r X_xtnd,  echo = FALSE, warning=FALSE}
X_xtnd <- extmin(X, relRef = rel_EX)
cat("Prior X extended in product space of (X,E","\n")
bcaPrint(X_xtnd)
```
Combine X extended and E_X in the product space $\Theta \times \Lambda$.
```{r relation2,  echo = FALSE, warning=FALSE}
comb_X_EX <- dsrwon(X_xtnd, rel_EX)
cat("Mass distribution of the combination of X extended and E_X","\n")
bcaPrint(comb_X_EX)
```
As we can see, we have

1. the distribution of the product $P(H) \cdot P(E|H)$ on $\Theta \times \Lambda$;

2. a mass allotted to the empty set $m(\varnothing)$, which is $1 - P(E)$.

Using function *nzdsr*, we apply the normalization constant to obtain the desired result. Then, using  function  *elim*, we obtain the marginal of X, which turns out to be $P(X | E = d)$

```{r relation3,  echo = FALSE, warning=FALSE}
norm_comb_X_EX <- nzdsr(comb_X_EX)
cat("The normalized mass distribution of the combination of X extended and E_X","\n")
bcaPrint(norm_comb_X_EX)
dist_XgE <- elim(norm_comb_X_EX, xnb = 4)
cat("The posterior distribution P(X|E) for (a,d), (b,d), (c,d), after eliminating variable E","\n")
bcaPrint(dist_XgE)
```








