---
title: "Bayes_Rule"
author: "Peiyuan Zhu"
date: "2023-11-02"
output: rmarkdown::html_vignette
# output: word_document
vignette: >
  %\VignetteIndexEntry{Bayes_Rule}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(dst) 
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In Mathematical Theory of Evidence Glenn Shafer talked about how Dempster's rule of combination generalizes Bayesian conditioning. In this document we investigate numerically how a simple Bayesian model can be encoded into the language of belief function. 

# Example

In particular, the Bayesian belief functions concentrates their masses on the singletons only, unlike more general basic mass assignment functions. For instance, in a frame $\Theta=\{a,b,c\}$, basic mass assignment $m(\{a\})=0.2$, $m(\{b\})=0.3$ and $m(\{c\})=0.5$ defines a Bayesian belief function. 

```{r}
Theta<-matrix(c(1,0,0,0,1,0,0,0,1,1,1,0,0,1,1,1,0,1,1,1,1), nrow = 7, byrow = TRUE)
bpa1 <- bca(tt=matrix(c(1,0,0,0,1,0,0,0,1), nrow = 3, byrow = TRUE), m = c(0.2, 0.3, 0.5), cnames = c("a", "b", "c"), idvar = 1)
belplau(bpa1, h=Theta)
```
The law of conditional probability is a special case of Dempster's rule of combination that all the masses focus on the event is conditioned. For instance, basic mass assignment focuses all the masses on subset $\{b,c\}$ so that $m(\{b,c\})=1$. 

```{r}
bpa2 <- bca(tt=matrix(c(0,1,1), nrow = 1, byrow = TRUE), m = c(1), cnames = c("a", "b", "c"), idvar = 2)
belplau(bpa2, h=Theta)
```
Combining the two basic mass assignments we get the belief after conditioning. Notice here prior and likelihood are symmetric.

```{r}
bpa3 <- dsrwon(bpa1, bpa2)
belplau(nzdsr(bpa3), h=Theta)
```
```{r}
bpa3 <- bca(tt=matrix(c(1,0,0,0,1,0,0,0,1), nrow = 3, byrow = TRUE), m = c(1, 0, 0), cnames = c("d", "e", "f"), idvar = 3)

tt <- matrix(c(1,0,0,1,0,0,
               1,0,0,0,1,0,
               1,0,0,0,0,1,
               0,1,0,1,0,0,
               0,1,0,0,1,0,
               0,1,0,0,0,1,
               0,0,1,1,0,0,
               0,0,1,0,1,0,
               0,0,1,0,0,1), ncol = 6, byrow = TRUE, dimnames = list(NULL, c("d", "e", "f", "a", "b", "c")))

spec <-  matrix(c(1,2,3,
                  4,5,6,
                  7,8,9,
                  0.1,0.2,0.7,
                  0.2,0.3,0.5,
                  0.4,0.5,0.1), ncol = 2, dimnames = list(NULL, c("specnb", "mass")))

inforvar <- matrix(c(1,3,3,3), ncol = 2,  dimnames = list(NULL, c("varnb", "size")) )

rel <- bcaRel(tt = tt, spec = spec, infovar = inforvar, varnames = c("v1", "v2"), relnb = 1)

bpa3_extnd<-extmin(bpa3,rel)
elim(rel,3)
```






