---
title: "Bayes_Rule"
author: "Peiyuan Zhu, Claude Boivin"
date: "2023-11-02"
output: rmarkdown::html_vignette
# output: word_document
vignette: >
  %\VignetteIndexEntry{Bayes_Rule}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# devtools::load_all(".") # only used in place of dst when testing with R-devel
library(dst) 
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In Mathematical Theory of Evidence Glenn Shafer talked about how Dempster's rule of combination generalizes Bayesian conditioning. In this document we investigate numerically how a simple Bayesian model can be encoded into the language of belief function. 

# Example

In particular, the Bayesian belief functions concentrates their masses on the singletons only, unlike more general basic mass assignment functions. For instance, in a frame $\Theta=\{a,b,c\}$, basic mass assignment $m(\{a\})=0.2$, $m(\{b\})=0.3$ and $m(\{c\})=0.5$ defines a Bayesian belief function. 

```{r bpa1 definition}
Theta<-matrix(c(1,0,0,0,1,0,0,0,1,1,1,0,0,1,1,1,0,1,1,1,1), nrow = 7, byrow = TRUE)
bpa1 <- bca(tt=matrix(c(1,0,0,0,1,0,0,0,1), nrow = 3, byrow = TRUE), m = c(0.2, 0.3, 0.5), cnames = c("a", "b", "c"), idvar = 1)
round(belplau(bpa1, h=Theta), digits = 3)
```
The law of conditional probability is a special case of Dempster's rule of combination that all the masses focus on the event is conditioned. For instance, basic mass assignment focuses all the masses on subset $\{b,c\}$ so that $m(\{b,c\})=1$. 

```{r bpa2 definition}
bpa2 <- bca(tt=matrix(c(0,1,1), nrow = 1, byrow = TRUE), m = c(1), cnames = c("a", "b", "c"), idvar = 2)
belplau(bpa2, h=Theta)
```
Combining the two basic mass assignments we get the belief after conditioning. Notice here prior and likelihood are symmetric.

```{r bpa3 definition}
bpa3 <- dsrwon(bpa1, bpa2)
round(belplau(nzdsr(bpa3), h=Theta), digits = 3)
```

```{r rellations}
bpa2 <- bca(tt=matrix(c(0,1,1), nrow = 1, byrow = TRUE), m = c(1), cnames = c("a", "b", "c"), idvar = 2, varnames = "X")
bpa4 <- bca(tt=matrix(c(1,0,0,0,1,0,0,0,1), nrow = 3, byrow = TRUE), m = c(1, 0, 0), cnames = c("d", "e", "f"), idvar = 4, varnames = "Y")
# 
# construction of a the first conditional probability distributions P(Y =d | X )
# P(Y = d | X = a) = 0.1
# P(Y = d | X = b) = 0.2
# P(Y = d | X = c) = 0.7

tt_d <- matrix(c(1,0,0,1,0,0,
               0,1,0,1,0,0,
               0,1,0,0,1,0,
               0,1,0,0,0,1,
               0,0,1,1,0,0,
               0,0,1,0,1,0,
               0,0,1,0,0,1,
               0,1,0,1,0,0,
               1,0,0,1,0,0,
               1,0,0,0,1,0,
               1,0,0,0,0,1,
               0,0,1,1,0,0,
               0,0,1,0,1,0,
               0,0,1,0,0,1,
               0,0,1,1,0,0,
               1,0,0,1,0,0,
               1,0,0,0,1,0,
               1,0,0,0,0,1,
               0,1,0,1,0,0,
               0,1,0,0,1,0,
               0,1,0,0,0,1,
               1,1,1,1,1,1
               ), ncol = 6, byrow = TRUE, dimnames = list(NULL, c("a", "b", "c", "d", "e", "f")))
rownames(tt_d) <- nameRows(tt_d)
spec_d <-  matrix(c(rep(1,7), rep(2,7), rep(3,7), 4, rep(0.1,7), rep(0.2,7), rep(0.7,7),0),
 ncol = 2, dimnames = list(NULL, c("specnb", "mass")))

inforvar_d <- matrix(c(2,4,3,3), ncol = 2,  dimnames = list(NULL, c("varnb", "size")) )

rel_d <- bcaRel(tt = tt_d, spec = spec_d, infovar = inforvar_d, varnames = c("X", "Y"), relnb = 1)
bcaPrint(rel_d)
#
# 
# construction of a the second conditional probability distributions P(Y =e | X )
# P(Y = e | X = a) = 0.2
# P(Y = e | X = b) = 0.3
# P(Y = e | X = c) = 0.5

tt_e <- matrix(c(1,0,0,0,1,0,
               0,1,0,1,0,0,
               0,1,0,0,1,0,
               0,1,0,0,0,1,
               0,0,1,1,0,0,
               0,0,1,0,1,0,
               0,0,1,0,0,1,
               0,1,0,0,1,0,
               1,0,0,1,0,0,
               1,0,0,0,1,0,
               1,0,0,0,0,1,
               0,0,1,1,0,0,
               0,0,1,0,1,0,
               0,0,1,0,0,1,
               0,0,1,0,1,0,
               1,0,0,1,0,0,
               1,0,0,0,1,0,
               1,0,0,0,0,1,
               0,1,0,1,0,0,
               0,1,0,0,1,0,
               0,1,0,0,0,1,
               1,1,1,1,1,1
               ), ncol = 6, byrow = TRUE, dimnames = list(NULL, c("a", "b", "c", "d", "e", "f")))
rownames(tt_e) <- nameRows(tt_e)
spec_e <-  matrix(c(rep(1,7), rep(2,7), rep(3,7), 4, rep(0.2,7), rep(0.3,7), rep(0.5,7),0),
 ncol = 2, dimnames = list(NULL, c("specnb", "mass")))

inforvar_e <- matrix(c(2,4,3,3), ncol = 2,  dimnames = list(NULL, c("varnb", "size")) )

rel_e <- bcaRel(tt = tt_e, spec = spec_e, infovar = inforvar_e, varnames = c("X", "Y"), relnb = 2)
bcaPrint(rel_e)
#
# construction of a the third conditional probability distributions P(Y =f | X )
# P(Y = f | X = a) = 0.4
# P(Y = f | X = b) = 0.5
# P(Y = f | X = c) = 0.1

tt_f <- matrix(c(1,0,0,0,0,1,
               0,1,0,1,0,0,
               0,1,0,0,1,0,
               0,1,0,0,0,1,
               0,0,1,1,0,0,
               0,0,1,0,1,0,
               0,0,1,0,0,1,
               0,1,0,0,0,1,
               1,0,0,1,0,0,
               1,0,0,0,1,0,
               1,0,0,0,0,1,
               0,0,1,1,0,0,
               0,0,1,0,1,0,
               0,0,1,0,0,1,
               0,0,1,0,0,1,
               1,0,0,1,0,0,
               1,0,0,0,1,0,
               1,0,0,0,0,1,
               0,1,0,1,0,0,
               0,1,0,0,1,0,
               0,1,0,0,0,1,
               1,1,1,1,1,1
               ), ncol = 6, byrow = TRUE, dimnames = list(NULL, c("a", "b", "c", "d", "e", "f")))
rownames(tt_f) <- nameRows(tt_f)
spec_f <-  matrix(c(rep(1,7), rep(2,7), rep(3,7), 4, rep(0.4,7), rep(0.5,7), rep(0.1,7),0),
 ncol = 2, dimnames = list(NULL, c("specnb", "mass")))

inforvar_f <- matrix(c(2,4,3,3), ncol = 2,  dimnames = list(NULL, c("varnb", "size")) )

rel_f <- bcaRel(tt = tt_f, spec = spec_f, infovar = inforvar_f, varnames = c("X", "Y"), relnb = 2)
bcaPrint(rel_f)
#
# Combine the three relations
comb1_rels_de <- dsrwon(rel_d, rel_e)
comb_rel_def <- dsrwon(comb1_rels_de, rel_f)
bcaPrint(comb_rel_def)
X_margin <- elim(comb_rel_def, xnb = 4)
bcaPrint(X_margin)
Y_margin <- elim(comb_rel_def, xnb = 2)
bcaPrint(Y_margin)
#
# Inference
bcaPrint(bpa4)
bpa4_extnd<-extmin(bpa4,comb_rel_def)
bcaPrint(bpa4_extnd)
comb_rels_def_bpa4 <- dsrwon(bpa4_extnd, comb_rel_def )
comb_rels_def_bpa4$con
# 0
bcaPrint(comb_rels_def_bpa4)
X_if_d <- elim(comb_rels_def_bpa4, xnb = 4)
bcaPrint(X_if_d)
round(belplau(X_if_d), digits=2)

```









