---
title: "Introduction to belief Functions: The Monty Hall Game"
author: "Claude Boivin, Stat.ASSQ"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{An example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction
In this vignette, I use a small problem, The Monty Hall Game, to show the use of some functions of the package. Hence, I will show:

* how to define a mass function *m* on some subsets of the set of possible values of a variable *A* (function *bca*);
* how to specify a relation *r* between three variables *A, B, C* by the way of a mass function *m* on some subsets of the set of possible values of the product space *AxBxC* (function *bcaRel*);
* how to extend a mass function to a larger space in order to combine it with another mass function  defined on this space (function *extmin*);
* how to combine two mass functions (functions *dsrwon* and *nzdsr*);
* how to eliminate a variable from a relation to obtain a reduced space (marginalization, with function elim.R);
* how to compute measures of belief and plausibility (functions *belplau* and *tabresul*).

## The Monty Hall Game

Let us recall the Monty Hall Game from its statement in the Wikipedia article on the subject ([Monty Hall problem](https://en.wikipedia.org/w/index.php?title=Monty_Hall_problem&oldid=829292640)):

    "Suppose you're on a game show. Three doors A, B and C are in front of you. Behind one door is a brand new car, and behind the two others, a goat. You are asked to pick one of the three doors. Then the host of the game, who knows what's behind the doors, opens one of the two remaining doors and shows a goat. He then asks to you: "Do you want to switch doors or keep your initial choice?" 

Say you have choosen door A and the host has opened door B.
The question now is: Is it to your advantage to switch your choice from door A to door C?"

**Some notation to begin with**

For each door *A, B, C*, consider the same frame _F_ with three possible values: 
  $$F = \{car, goat 1, goat 2\}.$$ 
I use boolean vectors to identify each element of the frame. Hence, the element "car"" is identified by the vector $(1, 0, 0)$, goat 1 by the vector $(0, 1, 0)$ and goat 2 by the vector $(0, 0, 1)$.

With this notation, any subset of *F* can has a unique boolean representation. For example, the subset $\{goat 1, goat 2\}$ is represented by the vector $(0, 1, 1)$.

## Analysis of the problem

We have three things to consider: 

1. how the three doors are linked;
2. the evidence pertaining to door A (choice of the  contestant);
3. the evidence pertaining to door B (the action of the host).

### 1. How the three doors are linked

There are six possible ways of disposing the car and the two goats behind the three doors *A, B, C*:
$\{car, goat 1, goat 2\}$
$\{car, goat 2, goat 1\}$
$\{goat 1, car, goat 2\}$
$\{goat 1, goat 2, car\}$
$\{goat 2, car, goat 1\}$
$\{goat 2, goat 1, car\}$.

This information, will be encoded in the product space *AxBxC*. The frame, or set of possible values *F~ABC~* of *AxBxC* is the product of the frames of *A, B* and *C*. Hence, its number of elements is $3^3 = 27$.  The six possible dispositions of the car and goats determine a subset *S* of the frame *F~ABC~*. A mass of 1 will be allotted to this subset *S*.

We use the function *bcaRel* to establish the desired relation between the doors. We prepare the necessary inputs and call the function.

```{r}
library(dst)
# 1. define the tt matrix MHABC_tt, which encodes the subset S
# 
MHABC_tt <- matrix(c(1,0,0,0,1,0,0,0,1,
                     1,0,0,0,0,1,0,1,0,
                     0,1,0,1,0,0,0,0,1,
                     0,1,0,0,0,1,1,0,0,
                     0,0,1,1,0,0,0,1,0,
                     0,0,1,0,1,0,1,0,0), ncol=9, byrow=TRUE)
colnames(MHABC_tt) <- rep(c("car", "g1", "g2"), 3)
#
# 2. define the spec matrix. 
# Here we have one subset of six elements
# 
MHABC_spec = matrix(rep(1,12), ncol = 2, dimnames = list(NULL, c("specnb", "mass"))) 
# 
# 3. define the info matrix. 
# for each variable, we attribute a number and give the size of the frame
# 
MHABC_info =matrix(c(1:3, rep(3,3)), ncol = 2, dimnames = list(NULL, c("varnb", "size")) )
# 
# 4. call of the function with the name of the variables and the numbering of the relation
# 
MHABC_rel <-  bcaRel(tt = MHABC_tt, spec = MHABC_spec, infovar = MHABC_info, infovarnames = c("MHA", "MHB", "MHC"), relnb = 1)
# 
# let's see some results:
# Note that row labels of the resulting tt matrix can become pretty long. If the case, we can use the prmatrix function, for example:
prmatrix(t(MHABC_rel$tt), collab = "")
# Another way to check the tt matrix is:
which(MHABC_rel$tt[1,] == TRUE)
```

### 2. the evidence pertaining to door A

You have chosen door A. At this point, the problem is quite simple. Your belief is equally divided between 3 possible outcomes: car, goat 1 or goat 2:
$m({car}) = m({g1}) = m({g2}) = 1/3$.
let's encode this evidence with function *bca*.
```{r}
# Evidence for door A
MHA_E <-  bca(f= diag(1,3,3), m= rep(1/3, 3), cnames =c("car", "goat1", "goat2"), infovarnames = "MHA", varnb = 1)
MHA_E
# At this point, no big deal...
tabresul(MHA_E)
```
### 3. the evidence pertaining to door B

But the host wanted to add some thrill to the game. He has opened door B and revealed a goat. The host has given us a small piece of evidence: Goat 1 or goat 2 was behind door B. Since the host knows what is behind each door, the mass value of this piece of evidence is:
$m({g1, g2}) = 1$.

Let's translate this in R with function *bca*:
```{r}
# Evidence for door B
MHB_E <- bca(f= matrix(c(0,1,1), ncol=3, byrow = TRUE), m=1, cnames =c("car", "goat1", "goat2"), infovarnames = "MHB" , varnb=2)
MHB_E
```
## The hypergraph of the Monty Hall game

We can now show The Monty Hall game as a belief network. Variables A, B and C are the nodes of the graph. The edges (hyperedges) are the evidences MHA_E AND MHB_E and the relation MHABC_rel. We use the package igraph ^[Csardi G, Nepusz T: The igraph software package for complex network research, InterJournal, Complex Systems 1695. 2006. http://igraph.org] to produce a bipartite graph corresponding to the desired hypergraph.

```{r, fig.show='hold', fig_caption: yes}
# The network
library(igraph)
# Encode the evidences and relations with an incidence matrix
Monty_hgm <- matrix(c(1,1,1,1,0,0,0,1,0), ncol=3, dimnames = list(c("A", "B", "C"), c("r_ABC", "ev_A", "ev_B")))
# The graph structure
Monty_hg <- graph_from_incidence_matrix(incidence = Monty_hgm, directed = FALSE, multiple = FALSE, weighted = NULL,add.names = NULL)
V(Monty_hg)
# Show variables as circles, relations and evidences as rectangles
V(Monty_hg)$shape <- c("circle", "crectangle")[V(Monty_hg)$type+1]
V(Monty_hg)$label.cex <- 0.6
V(Monty_hg)$label.font <- 2
# render graph
plot(Monty_hg, vertex.label = V(Monty_hg)$name, vertex.size=(4+4*V(Monty_hg)$type)*8)
```

## Now, the solution with the calculus of belief functions
To obtain the Belief function for door C, we have to combine evidences for doors A and B with the relation on doors ABC. We will proceed by elimination of the doors (variables), eliminating A first, then B.

### 1. Eliminate A from the network 
To do that, we extend Evidence on A to ABC, combine the extended evidence with the relation on ABC, then marginalize on BC. This gives a reduced network with B and C

```{r}
# Extend MHA to the product space AxBxC
MHA_ext <- extmin(MHA_E, MHABC_rel )
# Combine MHA_ext and MHABC_rel
MHA_ABC_comb <- dsrwon(MHA_ext,MHABC_rel, relnb = 2)
MHA_ABC_comb$con
# Eliminate variable A ( since  the measure of contradiction is 0, no need to normalize)
MHBC <- elim(MHA_ABC_comb, xnb = 1)
```
#### The reduced belief network
```{r}
Monty2_hgm <- matrix(c(1,1,1,0), ncol=2, dimnames = list(c("B", "C"), c("r_BC", "ev_B")))
Monty2_hg <- graph_from_incidence_matrix(incidence = Monty2_hgm, directed = FALSE, multiple = FALSE, weighted = NULL,add.names = NULL)
V(Monty2_hg)
# Variables as circles, relations and evidences as rectangles
V(Monty2_hg)$shape <- c("circle","crectangle")[V(Monty2_hg)$type+1]
V(Monty2_hg)$label.cex <- 0.6
V(Monty2_hg)$label.font <- 2
# render graph
# plot(Monty_hg, vertex.size=40)
plot(Monty2_hg, vertex.label = V(Monty2_hg)$name, vertex.size=(4+4*V(Monty2_hg)$type)*8)
```

### 2. Eliminate B from the network
Similarly, we extend Evidence on B to BC, combine the extended evidence with the relation on BC, then marginalize on C. This will give the final result.
```{r}
# Extend MHB_E to the space BxC
MHB_ext <- extmin(MHB_E, MHBC )
MHB_BC_comb <- dsrwon(MHB_ext, MHBC, relnb = 3)
MHB_BC_comb$con
# Eliminate variable B (since  MHA_BC_comb$con = 0, no need to normalize)
MHC <- elim(MHB_BC_comb, xnb = 2)
belplau(MHC)
```

## Conclusion:
As we can see we double our chances of winning the car if we swicth from door A to door C. 

Note that there is no loss of generality by fixing the choices in the analysis (door A for the contestant, door B for the host).

To be more specific and make a bridge with probability theory, we can add to our result all the elementary events that have 0 mass, so that we can see their measure of plausibility.

The function *addTobca* serves this purpose.
```{r}
MHC_plus_singl <- addTobca(MHC, f=matrix(c(0,1,0,0,0,1), ncol = 3, byrow = TRUE))
tabresul(MHC_plus_singl)
```
