---
title: "Reliability_Proof_Machinery"
author: "Peiyuan Zhu"
date: "2024-03-28"
output: rmarkdown::html_vignette
# output: word_document
vignette: >
  %\VignetteIndexEntry{Reliability_Proof_Machinery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# devtools::load_all(".") # only used in place of dst when testing with R-devel
knitr::opts_chunk$set(echo = TRUE)
```

Here's an example given in section 1.1 of Mathematical Theory of Hints by Jurg Kohlas and Paul-Andre Monney. Suppose we have implication $a_1 \vee a_2 \implies b$ while $a_1$, $a_2$ are not known to be true for certain. Let $p_1 = 0.3$ be the probability that $a_1$ is true and $p_2 = 0.4$ the probability that $a_2$ is true. This is an example of combining "pure arguments" by Jacob Bernoulli in Ars Conjectandi.

First, we use function *bcaRel* to define the implication relation in its disjunctive form $b \vee (\neg a_1 \land \neg a_2)$. The required binary table can also be obtained from https://web.stanford.edu/class/cs103/tools/truth-table-tool/.

```{r chk1}
tt <- matrix(c(0,1,0,1,0,1,
               1,0,1,0,1,0,
               1,0,1,0,0,1,
               0,1,1,0,0,1,
               1,0,0,1,0,1,
               1,1,1,1,1,1), nrow = 2 + 3 + 1, ncol = 6, byrow = TRUE, dimnames = list(NULL,c("a1 no", "a1 yes", "a2 no", "a2 yes", "b no", "b yes")))
spec <- matrix(c(1,1,1,1,1,2,1,1,1,1,1,0), nrow = 5 + 1, ncol = 2)
infovar <- matrix(c(1,2,3,2,2,2), nrow = 3, ncol = 2)
varnames <- c("a1","a2","b")
bcaRel1<-bcaRel(tt,spec,infovar,varnames)
cat("The implication relation","\n")
bcaPrint(bcaRel1)
```
Second, we use function *bca* to define the probabilities such that each of the assumptions are true. For $a_1$ is true, probability 0.3 is given to "a1 is true" and 0.7 given to the whole first frame.

```{r chk2}
tt <- matrix(c(0,1,1,1), nrow = 2, ncol = 2, dimnames = list(NULL, c("a1 no", "a1 yes")))
m <- c(0.3,0.7)
varnames <- "a1"
idvar <- 1
bca1 <- bca(tt, m, idvar=idvar, varnames=varnames)
bcaPrint(bca1)
```
For $a_2$ is true, probability 0.4 is given to "a2 is true" and 0.6 given to the whole second frame.
```{r chk3}
tt <- matrix(c(0,1,1,1), nrow = 2, ncol = 2, dimnames = list(NULL, c("a2 no", "a2 yes")))
m <- c(0.4,0.6)
varnames <- "a2"
idvar <- 2
bca2 <- bca(tt, m, idvar=idvar, varnames=varnames)
bcaPrint(bca2)
```
Now we combine the two bca's. To do that we need to first extend the two bca's are they're defined on the marginal frames. Using function *extmin*,  bca1 can be extended to the whole frame of the product space of the three variables (a1, a2, b) as:

```{r chk4}
bca1_extmin <- extmin(bca1,bcaRel1)
bcaPrint(bca1_extmin)
```
Likewise, bca2 can be extended to the whole frame of the product space as:
```{r chk5}
bca2_extmin <- extmin(bca2,bcaRel1)
bcaPrint(bca2_extmin)
```
Having extended the marginal bca to the whole frame, we can use function *dsrwon* to perform Dempster's rule of combination them. 

```{r chk6}
bca12_extmin <- dsrwon(bca1_extmin,bca2_extmin)
bcaPrint(bca12_extmin)
```

Remember that at the beginning, we defined a relation $a_1 \vee a_2 \implies b$. This relation must now be combined with the combined bca's to yield the final bca in the product space of (a1, a2, b).

```{r chk7}
bca12_extmin_dsrwon_bcaRel1 <- dsrwon(bca12_extmin,bcaRel1)
bcaPrint(bca12_extmin_dsrwon_bcaRel1)
```

Now we can get the marginal bca of variable *b*. To do so, we need to summarize the other variables on this dimension. We do so by eliminating (deleting) the other dimensions than *b*, that is *a1* and *a2*. We choose to eliminate dimension 1 (a1) first, using function *elim*. 
```{r chk8}
bca12_extmin_elim1 <- elim(bca12_extmin_dsrwon_bcaRel1,1)
bcaPrint(bca12_extmin_elim1)
```
Likewise, we eliminate dimension 2. 
```{r chk9}
bca12_extmin_elim12 <- elim(bca12_extmin_elim1,2)
bcaPrint(bca12_extmin_elim12)
```
Having obtained the marginal bca of variable *b*, we can now evaluate belief and plausibility, using function *belplau*. 

```{r chk10}
belplau(bca12_extmin_elim12)
```
Note the result: bel(yes) = 0.58;

which is the result one will obtain by applying the combination rule developed by Bernoulli:

$$bel(b) = 1 - (1-p1) \cdot (1-p2)$$
$$= 1 - (1-0.3 \cdot (1-0.4) = 1 - 0.42 = 0.58.$$
Alternatively, instead of using the OR gate, one can build up the graph by defining the two implications separately. First, we define the first implication.
```{r chk11}
tt <- matrix(c(0,1,0,1,
               1,0,0,1,
               1,0,1,0,
               1,1,1,1), nrow = 4, ncol = 4, byrow = TRUE, dimnames = list(NULL,c("a1 no", "a1 yes", "b no", "b yes")))
spec <- matrix(c(1,1,1,2,1,1,1,0), nrow = 4, ncol = 2)
infovar <- matrix(c(1,3,2,2), nrow = 2, ncol = 2)
varnames <- c("a1","b")
bcaRel1<-bcaRel(tt,spec,infovar,varnames)
bcaPrint(bcaRel1)
```

Similarly, we can define the second implication as follows.
```{r chk12}
tt <- matrix(c(0,1,0,1,
               1,0,0,1,
               1,0,1,0,
               1,1,1,1), nrow = 4, ncol = 4, byrow = TRUE, dimnames = list(NULL,c("a2 no", "a2 yes", "b no", "b yes")))
spec <- matrix(c(1,1,1,2,1,1,1,0), nrow = 4, ncol = 2)
infovar <- matrix(c(2,3,2,2), nrow = 2, ncol = 2)
varnames <- c("a2","b")
bcaRel2<-bcaRel(tt,spec,infovar,varnames)
bcaPrint(bcaRel2)
```

Then we extend, combine, and eliminate variables. 
For the first variable and the first implication, we obtain:
```{r chk13}
bca1_extmin <- extmin(bca1,bcaRel1)
bca1_extmin_bcaRel1_dsrwon <- dsrwon(bca1_extmin, bcaRel1)
bca1_extmin_bcaRel1_dsrwon_elim <- elim(bca1_extmin_bcaRel1_dsrwon, 1)
bcaPrint(bca1_extmin_bcaRel1_dsrwon_elim)
```
Similarly for the second variable and the second implication, we obtain:
```{r chk14}
bca2_extmin <- extmin(bca2,bcaRel2)
bca2_extmin_bcaRel2_dsrwon <- dsrwon(bca2_extmin, bcaRel2)
bca2_extmin_bcaRel2_dsrwon_elim <- elim(bca2_extmin_bcaRel2_dsrwon, 2)
bcaPrint(bca2_extmin_bcaRel2_dsrwon_elim)
```

Now, we combine the two results 
```{r chk15}
bca12 <- dsrwon(bca1_extmin_bcaRel1_dsrwon_elim,bca2_extmin_bcaRel2_dsrwon_elim)
bcaPrint(bca12)
```

Next, evaluate belief and plausibility.
```{r chk16}
belplau(bca12)
```


