---
title: "The Peter, John and Mary example of combination of belief functions"
author: Claude Boivin^[Retired Statistician, Stat.ASSQ]
date: 2019-04-09, revised 2026-04-18
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The original peter, John and Mary example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# devtools::load_all(".") # only used in place of dst when testing development version
library(dst)
```

## The facts^[Smets, P. (1988) Belief Functions. Non-Standard Logics for Automated Reasoning (P. Smets, A. Mandani, D. Dubois, H. Prade, ed.), Academic Press, New York, p. 254-286]

Mr Jones has been murdered. There are three suspects, Peter, John and Mary, and some witnesses.

Witness one is a janitor who was asleep just before the murder. He claims to have heard the victim scream, then seen a short man run out of the victim's house. 

Witness two is an old lady with a bad sighting, who lives across the street. She claims to have seen the crime from her window and states that the murderer was much taller than the victim.

Peter and John are men.
John and Mary are small.
Peter is a tall man, taller than the victim.

First, we show how to combine the facts using Dempster's rule. Second, we show how to represent this problem using a belief network.

## 1. The pooling of evidence

A state space is defined by the three suspects. Let's call it "Culprit".
$$Culprit = \{Peter, John, Mary\}$$
Function SSR is called to define the state space representation 
```{r state_space, echo = TRUE, warning=FALSE}
# cat("State space representation of the Culprit","\n" )
ss_Culprit <- SSR(varnames = "AMurder", idvar = 1, size = 3, cnames = c("Peter", "John", "Mary"))
#
```

After defining the culprit's SSR, we use function DSM to assess the pieces of evidence collected on the crime scene.

```{r Pooling1, echo = TRUE, warning=FALSE}
# cat("Witness1, A janitor asleep","\n")
witness1 <- DSM(tt = matrix(c(1,1,0,0,1,1,1,1,1), ncol=3, byrow=TRUE), m= c(0.5, 0.2, 0.3), ssr = ss_Culprit)
printDSM(witness1)
#
# cat("Witness2, An old lady with bad sighting saw a tall man","\n")
tall <- DSM(tt = matrix(c(1,0,0,1,1,1), ncol=3, byrow=TRUE), m= c(0.6, 0.4),ssr = ss_Culprit)
printDSM(tall)
```

### Combination by Dempster's Rule

We use function DSC combine evidences using Dempster's Rule. Witness1 says a man (m = 0.5), small (m = 0.2). Witness2 says a tall man (Peter, m = 0.6). There is a bit of contradiction between witnesses. Hence function normalize will be called on the result of DSC.

```{r Pooling2, echo = FALSE, warning=FALSE}
# w_and_tall <- DSC(witness1_plus_singl, tall)
w_and_tall <- DSC(witness1, tall)
norm_w_and_tall <- normalize(w_and_tall)
result <- tabresul(norm_w_and_tall)
cat("Combination of witnesses 1 and 2","\n")
round(result$mbp, digits = 2)
cat("Conflict between evidence","\n")
result$Conflict

```
We observe that Peter's plausibility ratio is 2 compared to John or Marie. No fact directly points to John or Marie. We would like to know John's plausibility ratio, as well as Marie's. We will therefore add these hypotheses to the result with a mass of zero.
```{r Pooling999, echo = TRUE, warning=FALSE}
#
cat("Add all singletons in order to compare them","\n")
norm_w_and_tall_plus_singl <- addToDSM(norm_w_and_tall, tt = diag(3))
result <- tabresul(norm_w_and_tall_plus_singl)
round(result$mbp, digits = 2)
```
Peter (a tall man) is the more plausible culprit (belief of 0.55 with plausibility ratio of 2). This is the one for which the two pieces of evidence are the less contradictory. John and Mary are small, Mary is a woman They have each one a belief score of 0, since no evidence points exactly toward them. 

### Adding a third piece of evidence
Witness 3: Peter's girlfriend testifies that Peter was at her home far away from the victim's house when the crime happened. Peter's girlfriend says Peter was with him all night. We use function DSC to apply Dempster's Rule and combine her testimony with the previous results.

```{r Pooling3, echo = TRUE, warning=FALSE}
girlfriend <- DSM(tt = matrix(c(0,1,1,1,1,1), ncol=3, byrow=TRUE), m= c(1, 0), ssr = ss_Culprit)
cat("Witness 3, the girlfriend clears Peter","\n")
printDSM(girlfriend)
girlfriend_w_and_tall <- DSC(norm_w_and_tall_plus_singl, girlfriend)
norm_girlfriend_w_and_tall <- normalize(girlfriend_w_and_tall)
result <- tabresul(norm_girlfriend_w_and_tall)
round(result$mbp, digits = 3)
#
```
If we accept the girlfriend's testimony, Peter is now exonerated. However, if there is any doubt, the likelihood ratio between Pierre and John or Marie will vary from 0 to 2, depending on the degree of doubt placed on the girlfriend's testimony.

You can test the effect of the girlfriend's testimony by changing the weighting assigned to it and observe its influence on the plausibility of Pierre's guilt.

## 2. Viewing the problem as a belief network

### Facts and relations

Instead of grouping the three suspects Peter, John and Marie into the same state space, we consider that each suspect has their own state space: guilty (yes) or not guilty (no).

Peter:
$$P = \{yes, no\}$$
John:
$$J = \{yes, no\}$$
Mary:
$$M = \{yes, no\}$$
The testimony of a witness *k* is represented by a Dempster-Shafer Model
$$DSM_k(tt_k,m_k),$$
where $tt_k$ is the table of states induced by $witness_k$ in the product space $F_{PJM}$ and $m_k$ the vector of masses associated to the states.

The combination of evidences is summarized by the equation
$$ DSM_{F_{PJM}}(tt, m) = \bigoplus_{k=1}^{n} DSM_k(tt_k,m_k).$$

All evidences can be combined in the state space representation given by the product space $F_{PJM} = \prod(P, J, M)$, provided that they are defined in that space. If two evidences are not defined on the same space, they can be extended appropriately to the state space representation *F* to make them combinable. Let's see how it works.

Evidence of witness 1 is defined in the product space $F_{PJM} = \prod(P, J, M)$. Here is the code for witness 1.

```{r network1a, echo = TRUE, warning=FALSE}
# cat(" Define state space representations of marginals P, J, M","\n")
ss_P <- SSR(varnames = "Peter", idvar = 1, size = 2, cnames = c("yes", "no"))
ss_J <- SSR(varnames = "John", idvar = 2, size = 2, cnames = c("yes", "no"))
ss_M <- SSR(varnames = "Mary", idvar = 3, size = 2, cnames = c("yes", "no"))

# cat("witness 1 induced relation ","\n")
# cat(" Define state space representation F_PJM","\n")
ssr_PJM <- list(ss_P, ss_J, ss_M)
# cat("The description matrix of the relation","\n")
witness1_tt <- matrix(c(1,0,0,1,0,1,
                        0,1,1,0,0,1,
                        0,1,1,0,0,1,
                        0,1,0,1,1,0,
                        1,1,1,1,1, 1), ncol=6, byrow=TRUE)
colnames(witness1_tt) <- rep(c("yes", "no"), 3 )
witness1_tt
witness1_spec <- matrix(c(1,1,2,2,3, rep(0.5, 2), rep(0.2, 2), 0.3), ncol = 2, dimnames = list(NULL, c("specnb", "mass")))

witness1_rel <- jointDSM(tt = witness1_tt, spec = witness1_spec, ssrlist = ssr_PJM, relnb = 1)
# cat(" the testimony Peter or John (0.5), John or Mary (0.2) in the product space","\n")
printDSM(witness1_rel)
```
Evidence of witness 2 is defined in the space $F_{P}$.

```{r network1b, echo = TRUE, warning=FALSE}
#
# cat("Evidence for witness 2","\n")
witness2 <- DSM(tt = matrix(c(1,0,1,1), ncol=2, byrow=TRUE), m= c(0.6, 0.4), ssr = ss_P)
# cat("Peter (.6)","\n")
printDSM(witness2)
```
Evidence of witness 3 is defined in the product space $F_{JM} = \prod(J, M)$.

```{r network1c, echo = TRUE, warning=FALSE}
# cat("witness 3 induced relation ","\n")
# cat("The description matrix of the relation","\n")
witness3_tt <- matrix(c(0,1,1,0,
                        1,0,0,1,
                        1,1,1, 1), ncol=4, byrow=TRUE)
colnames(witness3_tt) <- rep(c("yes", "no"), 2 )
witness3_tt
witness3_spec <- matrix(c(1,1,2, 1, 1, 0), ncol = 2, dimnames = list(NULL, c("specnb", "mass")))
witness3_rel <- jointDSM(tt = witness3_tt, spec = witness3_spec, ssrlist = list(ss_J, ss_M), relnb = 2)
# cat("The testimony John or Mary (1) expressed in the product space","\n")
printDSM(witness3_rel)
```
Finally, a constraint on the product space  $F_{PJM}$ must be added to reflect the fact that there can only be one culprit. For example, we want to obtain the same result if we put $Peter = \{no\}$ instead of $(John, Mary) =\{\{yes, no\}, \{no, yes\}\}$ for witness 3.Hence, we put a mass of 1 on the subset 

$C = \{Peter, ¬John, ¬Mary\}\cup \{¬Peter, John, ¬Mary\}\cup \{¬Peter, ¬John, Mary\}$ ("¬" means "negation"). 

```{r network1d, echo = TRUE, warning=FALSE}
#
# cat("Adding a constraint on the state space F_PJM","\n")
# 1. The tt table
tt_PJM <- matrix(c(1,0,0,1,0,1,
                   0,1,1,0,0,1,
                   0,1,0,1,1,0,
                   rep(1,6)),
  nrow = 4, byrow = TRUE, 
  dimnames = list(NULL, rep(c("yes", "no"), 3 ) ) )
#  
# 2. The mass distribution
spec_PJM <-  matrix(c(rep(1,6),2, 0) , ncol = 2, byrow = TRUE,
dimnames = list(NULL, c("specnb", "mass")))
#
PJM_rel <- jointDSM(tt = tt_PJM, spec = spec_PJM, ssrlist = ssr_PJM, relnb = 3)
# cat("Linking the variables","\n")
# cat("Detective's assumption expressed in the product space","\n")
printDSM(PJM_rel)
#
```

Finally we have a state space which is the combination 3 variables, one piece of evidence against Peter, one evidence expressed in the product space  $F_{PJM}$ the other in the product space  $F_{JM}$ and a constraint in $F_{PJM}$ saying there can be only one culprit.

We construct the hypergraph, use function *uproj* to extend all the relations in the product space $\prod(P,J, M)$ combine the relations with function *DSC*, then use function *dproj* to obtain the three marginals.

### The Hypergraph

```{r, fig.show='hold', fig_caption: yes, echo=FALSE, message=FALSE}
# The network
if (requireNamespace("igraph", quietly = TRUE) ) {
library(igraph)
cat("Encode pieces of evidence and relations with an incidence matrix","\n")
N <- 3 # 3 variables
r_W1 <- 1*1:N %in% witness1_rel$infovar[,1]
ev_W2 <- 1*1:N %in% witness2$infovar[,1]
r_W3 <- 1*1:N %in% witness3_rel$infovar[,1]
r_Detective <- 1*1:N %in% PJM_rel$infovar[,1]
#
# the incidence matrix
PJM_hgm <- matrix(c(r_W1, ev_W2,r_W3 ,r_Detective), ncol=4, dimnames = list(c("Peter", "John", "Mary"), c("r_W1", "ev_W2", "r_W3", "r_Detective")))
cat("The incidence matrix of the Hypergraph","\n")
print(PJM_hgm)
#
cat("information on variables necessary for the Peeling algorithm","\n")
PJM_vars1 <- c(PJM_rel$valuenames)
PJM_vars <- rbind(PJM_rel$infovar)
PJM_var_names <-names(PJM_vars1)
rownames(PJM_vars) <- PJM_var_names
#
# infos on relations
PJM_data_names <- c("witness1_rel", "witness2", "witness3_rel", "PJM_rel")
PJM <- list(PJM_hgm, PJM_var_names, PJM_data_names)
print(PJM)
#
## The graph structure of the problem
#
PJM_hg <- graph_from_biadjacency_matrix(incidence = PJM_hgm, directed = FALSE, multiple = FALSE, weighted = NULL,add.names = NULL)
V(PJM_hg)
# Show variables as circles, relations and evidence as rectangles
V(PJM_hg)$shape <- c("circle", "crectangle")[V(PJM_hg)$type+1]
V(PJM_hg)$label.cex <- 0.6
V(PJM_hg)$label.font <- 2
# render graph
plot(PJM_hg, vertex.label = V(PJM_hg)$name, vertex.size=(3+6*V(PJM_hg)$type)*6, sub="Belief network for Peter,John, Mary's Example")
##
#
cat("\n")
}
```

### Study of extensions and combinations
```{r Combining the evidence, echo = FALSE, warning=FALSE}
cat("Combining evidence in the product space P x J x M","\n")
cat("We use the constraint PJM_rel as a reference to extend the others.", "\n")
#
cat("Extend Witness2 on PJM","\n")
w2_xtnd <- uproj(witness2, relRef = PJM_rel)
printDSM(w2_xtnd)
cat("Extend Witness 3 on PJM","\n")
w3_xtnd <- uproj(witness3_rel, relRef = PJM_rel)
printDSM(w3_xtnd)
#
cat("Combining the four relations on PJM","\n")
cat("Combining PJM_rel with witness1_rel","\n")
rel_comb1 <- DSC(PJM_rel,witness1_rel)
cat("rel_comb1", "\n")
printDSM(rel_comb1)

cat("Combining rel_comb1 with w2_xtnd","\n")
rel_comb2 <- DSC(rel_comb1, w2_xtnd)
cat("rel_comb2", "\n")
printDSM(rel_comb2)
#
cat("Combining rel_comb2 with w3_xtnd","\n")
rel_comb3 <- DSC(rel_comb2,w3_xtnd )
cat("rel_comb3", "\n")
printDSM(rel_comb3)
#
cat("Redo the combinations with a modifiied representation of witness 3","\n")
cat("Define w3 par ¬Peter (1), instead of jojn or Mary","\n")
#
witness3b <- DSM(tt = matrix(c(0,1,1,1), ncol=2, byrow=TRUE), m= c(1, 0), cnames =c("yes", "no"), varnames = "Peter", idvar = 1)
printDSM(witness3b)
#
cat("Extend Witness 3 on PJM","\n")
w3b_xtnd <- uproj(witness3b, relRef = PJM_rel)
printDSM(w3b_xtnd)
#
cat("combining with w3b_xtnd.  ")
cat("Note the differennce with the preceding result","\n")
rel_comb3b <- DSC(rel_comb2,w3b_xtnd )
printDSM(rel_comb3b)

# cat("combining with w3b_xtnd.  ")
# cat("Note the differennce with the preceding result","\n")
# rel_comb2b <- DSC(rel_comb,w3b_xtnd )
# printDSM(rel_comb2b)
cat("Normalizing...","\n")
# rel_comb3b_norm <- normalize(rel_comb3b)
# printDSM(rel_comb3b_norm)
# rel_comb3b_norm$con
#
# cat("Combining rel_comb2 rith PJM_rel","\n")
# rel_comb3 <- DSC(rel_comb2,PJM_rel )
# printDSM(rel_comb3)
#
# cat("Combining rel_comb2b rith PJM_rel","\n")
# rel_comb3b <- DSC(rel_comb2b,PJM_rel )
# printDSM(rel_comb3b)
cat("Now we obtain the same result","\n")
# rel_comb3_norm <- dst::normalize(rel_comb3)
# printDSM(rel_comb3_norm)
# rel_comb3_norm$con
# Test
rel_comb3b_norm <- dst::normalize(rel_comb3b)
printDSM(rel_comb3b_norm)
rel_comb3b_norm$con
# end test
#
cat("Compute marginals","\n")
cat("Peter","\n")
m_PJ <- dproj(rel_comb3b_norm, xnb = 3)
m_P <- dproj(m_PJ, xnb = 2)
printDSM(m_P)
cat("John or Mary","\n")
m_JM <- dproj(rel_comb3b_norm, xnb = 1)
printDSM(m_JM)
cat("John","\n")
m_J <- dproj(m_JM, xnb = 3)
printDSM(m_J)
cat("Mary","\n")
m_M <- dproj(m_JM, xnb = 2)
printDSM(m_M)
#
# cat("Combining rel_comb2 rith PJM_rel","\n")
# rel_comb3 <- DSC(rel_comb2,PJM_rel )
# printDSM(rel_comb3)
# rel_comb3_norm <- normalize(rel_comb3)
# printDSM(rel_comb3_norm)
# rel_comb3_norm$con
# #
# cat("Combining rel_comb2b rith PJM_rel","\n")
# rel_comb3b <- DSC(rel_comb2b,PJM_rel )
# printDSM(rel_comb3b)
# rel_comb3b_norm <- normalize(rel_comb3b)
# printDSM(rel_comb3b_norm)
# rel_comb3b_norm$con
# #
# cat("Compare rel_comb3 and rel_comb3b")
# all.equal(rel_comb3_norm,rel_comb3b_norm)
```
