---
title: "Using dst Package on Zadeh's Example"
author: "Claude Boivin, Stat.ASSQ"
date: "2020-01-19"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using dst Package on Zadeh's Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
#
# devtools::load_all(".") # only used in place of dst when testing with R-devel
# attach package dst
library(dst)
#
# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The criticism of Dempster-Shafer Theory (DST) by L. A. Zadeh ^[L. A. Zadeh. A mathematical theory of evidence (book review). AI Magazine, 55(81—83), 1984] has generated a lot of discussions and articles on the subject of conflicting evidence. In 2005, R. Haenni ^[R. Haenni. Shedding New Light on Zadeh’s Criticism of Dempster’s Rule of Combination. Conference: Information Fusion, 2005 8th International Conference] showed that the surprising result obtained by applying Dempster's rule to Zadeh's example was more due to the modelling of the situation than Dempster's rule not working. I add my grain of salt to the debate on Zadeh's example by showing a formulation of the problem as a small belief network and using Dempster's rule of combination to obtain a realistic result. This belief network gives the same results as the combination of the two evidences with the disjunctive rule of combination ^[P. Smets (1993). Belief Functions: The Disjunctive Rule of Combination and the Generalized Bayesian Theorem. IRIDIA - Université Libre de Bruxelles, Brussels, Belgium]. At the same time, I show how to use my *R* package *dst* ^[https://cran.r-project.org/package=dst] to do calculations on a belief network.

## Zadeh's Example
We suppose that a patient is examined by two doctors, A and B. A’s diagnosis is that P has either meningitis (M), with probability 0.99, or brain tumor (T), with probability 0.01. B agrees with A that the probability of a brain tumor is 0.01, but believes that it is the probability of concussion (C) rather than meningitis that is 0.99. 

Zadeh considers the same space of diseases {M, T, C} for the two experts. Hence, after the combination of the two pieces of evidence by Dempster's rule, we find as a result that the belief of a brain tumor is certain. 

```{r, echo = FALSE, warning=FALSE}
# Diagnosis from Expert 1. Coding the evidence with the bca function
ZExpert1 <- bca(f= matrix(c(1,0,0,0,1,0,1,1,1), ncol=3, byrow=TRUE), m= c(0.99, 0.01, 0), cnames =c("M", "T", "C"), infovarnames = "Diagnosis1", varnb = 1)
# show the definition of ZExpert1
cat("Diagnosis of Expert 1")
bcaPrint(ZExpert1)
#
# Diagnosis from Expert 2. Coding the evidence with the bca function
ZExpert2 <- bca(f= matrix(c(0,1,0,0,0,1,1,1,1), ncol=3, byrow=TRUE), m= c(0.01, 0.99, 0), cnames =c("M", "T", "C"), infovarnames = "Diagnosis2", varnb = 2)
# show the definition of ZExpert2
cat("Diagnosis of Expert 2")
bcaPrint(ZExpert2)
# Combination of Expert 1 and Expert 2 using Dempster's rule
cat("Combination of the two experts by Dempster's rule")
Ze1e2 <- nzdsr(dsrwon(ZExpert1, ZExpert2, relnb = 1))
tabresul(Ze1e2)
```

Indeed, the result does not reflect the opinions of the two experts. Before rejecting Dempster's rule as inappropriate to this situation, let's look more closely at the problem at hand. 

# A correct solution using Dempster's rule of combination

## Diagnosis of the two experts

Let's take *Expert one*.  *Expert one* distributes the whole mass between the two singletons {M} and {T}. *Expert one* does not consider {C} as a possibility. Hence we conclude that the Fod of *Expert one* cannot be {M, T, C}. We say that *Expert one* has restricted the space of possibilities of his/her diagnostic to the Fod {M, T}:

$Fod(D1) = \{M, T\}$. For simplicity, we write $D1 =\{M, T\}$.

The same line of reasoning is applied to *Expert two*. The whole mass of one is allotted to the set {T, C}, and the third possibility (M) is not considered at all.
Hence $D2 = \{T, C\}$.

I show the coding of these two pieces of evidence with the function *bca* of the package *dst*. 
```{r}
# attach package dst
library(dst)
#
# Diagnosis from first expert (function e1 attached to variable D1)
e1 <- bca(f= matrix(c(1,0,0,1,1,1), ncol=2, byrow=TRUE), m= c(0.99, 0.01, 0), cnames =c("M", "T"), infovarnames = "D1", varnb = 1)
#
# show the definition of e1
# "Diagnosis of Expert 1 (function e1 attached to variable D1)"
bcaPrint(e1)
#
# Diagnosis from second expert (function e2 attached to variable D2)
e2 <- bca(f= matrix(c(1,0,0,1,1,1), ncol=2, byrow=TRUE), m= c(0.99, 0.01, 0), cnames =c("C", "T"), infovarnames = "D2", varnb = 2)
#
# show the definition of e2
# "Diagnosis of Expert 2 (function e2 attached to variable D2)"
bcaPrint(e2)
```

## Linking the Experts to... the Patient
The two experts are reasoning in two different spaces of possibilities. To be able to combine their diagnosis, we need a common ground. This can be done if we introduce a third person, the patient, with a variable of interest, his/her disease (D). Then it is natural to take the union of the Fod of *Expert one* and *Expert two* as the Fod of the patient:

  $D = \{M, T\} \cup \{C, T\} = \{M, T, C\}$.

Hence, the patients' evaluation of his/her disease involves the disjunction of the evidences of the two experts.This situation is described by a relation of implication between experts and patient:

r1: $D1 \cup D2 \rightarrow D$.

The relation *r1* is represented in the product space $\prod(D1, D2, D)$ by one focal set of mass one:

$m(M C M + M C C + M T M + M T T + T C T + T C C + T T T) = 1$ (for simplicity, the "+" sign is used as the $\vee$ disjunctive operator in the functions of the package dst). Now I use the function *bcaRel* to code this relation.

```{r}
# 1. Defining the relation with a (0,1) matrix
tt_r1 <- matrix(c(1,0,1,0,1,0,0,1,0,1,0,0,0,1,1,0,0,1,1,0,0,1,0,0,1,0,1,0,0,1,1,0,0,1,0,0,1,1,0,0,0,1,0,1,0,1,0,1,0,1,1,1,1,1,1,1), ncol = 7,byrow = TRUE)
colnames(tt_r1) = c("M", "T", "C", "T", "M", "T", "C")
#
# 2. Setting the mass function
spec_r1 = matrix(c(rep(1,7),2, rep(1,7), 0), ncol = 2, dimnames = list(NULL, c("specnb", "mass"))) 
#
# 3. Names of variables names and dimension of their Fod
info_r1 =matrix(c(1:3, 2,2,3), ncol = 2, dimnames = list(NULL, c("varnb", "size")) )
#
#  The relation between e1, e2 and a patient p
r1 <-bcaRel(tt = tt_r1, spec = spec_r1, infovar = info_r1, infovarnames = c("D1", "D2", "D"), relnb = 1)
# show the relation
# "Relation between Experts and patient defined in the product space D1 x D2 x D"
bcaPrint(r1)
```

## The belief network for this situation
We now have all the elements of a small network made of one relation (r1) between three variables: Disease (D), Diagnosis1 (D1), Diagnosis2 (D2),  and two pieces of evidence coming from *Expert one* (e1) and *Expert two* (e2).

The three variables are the nodes of the graph: Disease (D), Diagnosis1 (D1), Diagnosis2 (D2). The edges (hyperedges) are given by the relation r1 and the two pieces of evidence (*Expert one* and *Expert two*). Using the igraph package, ^[Csardi G, Nepusz T: The igraph software package for complex network research, InterJournal, Complex Systems 1695. 2006. http://igraph.org] a bipartite graph corresponding to the hypergraph is obtained.

```{r, fig.show='hold', fig_caption: yes, echo=FALSE, message=FALSE}
# The network
if (requireNamespace("igraph", quietly = TRUE) ) {
library(igraph)
# Encode pieces of evidence and relations with an incidence matrix
rel1 <- 1*1:3 %in% r1$infovar[,1]
ev1 <- 1*1:3 %in% e1$infovar[,1]
ev2 <- 1*1:3 %in% e2$infovar[,1]

# information on variables
meddiag_vars1 <- c(r1$infovaluenames)
meddiag_vars <- rbind(r1$infovar)
meddiag_var_names <-names(meddiag_vars1)
rownames(meddiag_vars) <- meddiag_var_names
# infos on relations
meddiag_data_names <- c("e1", "e2", "r1")
# the incidence matrix
meddiag_hgm <- matrix(c(ev1,ev2, rel1), ncol=3, dimnames = list(c("D1", "D2", "D"), c("e1","e2", "r1")))
meddiag <- list(meddiag_hgm, meddiag_var_names, meddiag_data_names)
#
## The graph structure of the problem
#
meddiag_hg <- graph_from_incidence_matrix(incidence = meddiag_hgm, directed = FALSE, multiple = FALSE, weighted = NULL,add.names = NULL)
V(meddiag_hg)
# Show variables as circles, relations and evidence as rectangles
V(meddiag_hg)$shape <- c("circle", "crectangle")[V(meddiag_hg)$type+1]
V(meddiag_hg)$label.cex <- 0.6
V(meddiag_hg)$label.font <- 2
# render graph
plot(meddiag_hg, vertex.label = V(meddiag_hg)$name, vertex.size=(3+6*V(meddiag_hg)$type)*6, sub="Hypergraph (Belief network) for Zadeh's Example")
}
```

## Calculations on the Belief Network

Our goal is the calculation of a belief function "p" attached to the variable of interest "D" (Disease of the patient). At the beginning of the process, this function is simply the vacuous specification:

$p(\{M, T, C\}) = 1$.

We proceed by successive elimination of variables until only  the variable *D* remains, and the result of the function *p* is obtained.

The calculations involved are following the principles of the valuation language of Shenoy ^[P. P. Shenoy. A Valuation-Based Language for Expert systems. lnternational Journal of Approximate Reasoning 1989, 3 383--411]; see also ^[P. P. Shenoy. Valuation-Based Systems. Third School on Belief Functions and Their Applications, Stella Plage, France. September 30, 2015]. The variables are linked to functions (called valuations). A function can be a piece of evidence attached to a variable or a relation between two or more variables.

Three kinds of operations are involved in the calculations:
a) the minimal (vacuous) extension of a mass function to a larger Fod;
b) the combination of two mass functions by Dempster's rule;
c) the marginalization of a mass function, i.e. eliminating a variable to reduce the function to a smaller Fod. Let's do it.

First step: eliminate variable D1 (Diagnosis1). Using function *extmin*, we extend the mass function e1 to the space $\prod(D1, D2, D)$; then we combine e1 extended with r1, using functions *dsrwon* and *nzdsr* (normalization); finally, we use function *elim* to eliminate D1 by marginalizing to $\prod(D2, D)$. The mass function obtained is named rel_2.

```{r}
# some settings to begin
# variables numbers: D1=1, D2=2, D=3
N <- 1:3
# Elimination order of the variables. The goal: Disease (variable 3)
elim_order <- c(1,2,3)
var_to_elim <- rownames(meddiag_hgm)[order(elim_order)]
#
# 1: first step
# "first variable to eliminate"
cat(var_to_elim[1]) # Diagnosis1 (1)
irel_to_elim<- meddiag_hgm["D1",]*1:ncol(meddiag_hgm)
rels_nb <- irel_to_elim[irel_to_elim>0]
#
# 1 Fusion of Expert1 and Relation 1 and eliminate variable Diagnosis1
# 1.1 Extension of e1 (Expert1)
Expert1_ext <- extmin(get(meddiag_data_names[1]), get(meddiag_data_names[3])) 
# "Evidence of Expert 1 extended to the product space D1 x D2 x D"
bcaPrint(Expert1_ext)
#
# 1.2 Combination of Expert1 and r1
r2 <- nzdsr(dsrwon(Expert1_ext, get(meddiag_data_names[3]), relnb = 1 + length(meddiag_data_names)))
# "Subsets resulting from the combination of Expert 1 extended and r1"
bcaPrint(r2)
#
# 1.3 Marginalization (eliminate variable D1)
rel_2 <- elim(r2, xnb = order(elim_order)[1])
# "subsets in the product space D2 x D (after elimination of D1"
bcaPrint(rel_2)
```
After this first step, the graph is updated.

```{r echo=FALSE}
# update hg and relations names
# remove e1 and r1, add r2
R2 <- 1*1:3 %in% rel_2$infovar[,1]
meddiag_hgm1 <- cbind(meddiag_hgm[,-c(1,3)], R2)
colnames(meddiag_hgm1) <- c("e2", "r2")  
meddiag_data_names1 <- c(meddiag_data_names[-c(1,3)], "rel_2")
meddiag_var_names1 <- meddiag_var_names[-order(elim_order)[1]]
meddiag <- list(meddiag_hgm1, meddiag_var_names1, meddiag_data_names1)
meddiag_hg <- graph_from_incidence_matrix(incidence = meddiag_hgm1, directed = FALSE, multiple = FALSE, weighted = NULL,add.names = NULL)
# V(meddiag_hg)
# Show variables as circles, relations and evidence as rectangles
V(meddiag_hg)$shape <- c("circle", "crectangle")[V(meddiag_hg)$type+1]
V(meddiag_hg)$label.cex <- 0.6
V(meddiag_hg)$label.font <- 2
# render graph
plot(meddiag_hg, vertex.label = V(meddiag_hg)$name, vertex.size=(3+6*V(meddiag_hg)$type)*6, sub = "Reduced belief network after elimination of variable D1")
```

Second step: Eliminate variable D2 (Diagnosis2). We extend e2 to the space $\prod(D2, D)$; we combine e2 extended with rel_2 and marginalize to D to get the final result.

```{r}
# 2. Second step
# Fusion of Expert2 and Relation 2 and eliminate diagnosis2
#
# 2.1 Extension of Expert2 (e2)
# here, use rel_2 as the relation of reference for extmin 
Expert2_ext <- extmin(e2, rel_2) 
temp1 <- as.data.frame(cbind(rownames(Expert2_ext$tt), Expert2_ext$spec))
colnames(temp1)[1] <- "Expert2_ext"
# "Evidence of Expert 2 extended to the product space D2 x D"
print(temp1)
#
# 2.2 combinaison pf Expert2 and rel_2
r3 <- nzdsr(dsrwon(Expert2_ext, rel_2, relnb = 3))
temp1 <- as.data.frame(cbind(rownames(r3$tt), r3$spec))
colnames(temp1)[1] <- "r3"
# "Subsets of the space D2 x D resulting from the combination of Expert 2 extended and r2"
print(temp1)
#
# 2.3 Final result: the belief function p attached to variable D)
p <- elim(r3, xnb = order(elim_order)[2])
#
# add singletons with 0 mass to show all singletons in the results
p_sing <- addTobca(p,  f = matrix(c(1,0,0,0,0,1), ncol=3))
# "The final result after elimination of variable D2"
tabresul(p_sing)
```

The plausibility ratio column shows that the odds for $M \vee C$ against T are 50:1.  hence, the disease of the patient must be M or C. We also see that each single hypothesis, M, or, C, remains highly plausible (0.99). Although there is some support for T, its plausibility is very weak at 0.019. Finally, we use the plausibility transformation ^[Cobb, B. R. and Shenoy, P.P. (2006). On the plausibility transformation method for translating belief function models to probability models. Journal of Approximate Reasoning, 41(3), April 2006, 314–330] to look at the probability distribution over the Fod {M, T, C} of the patient that we obtain

```{r}
plautrans(p)
```
Now we have very high odds for M against T or for C against T: 495:1.

